### 2020-06-04 README.Rmd

```{r, eval = run_bw}
# Read in Big Woods census data from 2008 & 2014
bw_2008 <- 
  "https://deepblue.lib.umich.edu/data/downloads/z603qx485" %>% 
  read_delim(delim = "\t") %>% 
  mutate(spcode = to_any_case(spcode)) %>%
  select(
    treeID = treeid, stemID = stemtag, sp = spcode, quadrat, gx, gy, dbh, 
    date, codes
  )
bw_2014 <- 
  "https://deepblue.lib.umich.edu/data/downloads/1831ck00f" %>% 
  read_delim(delim = "\t") %>% 
  mutate(spcode = to_any_case(spcode)) %>%
  select(
    treeID = treeid, stemID = stemtag, sp = spcode, quadrat, gx, gy, dbh, 
    date, codes
  )

# Read in bigwoods grouping classification data
bw_species <- 
  "https://deepblue.lib.umich.edu/data/downloads/000000086" %>% 
  read_delim(delim = "\t") %>% 
  # convert all to snake case:
  mutate_at(c("species", "genus", "family", "idlevel", "spcode"), to_any_case) %>% 
  # join trait group
  left_join(families, by = c("spcode", "family")) %>% 
  mutate(
    sp = str_sub(genus, 1, 2), 
    sp = str_c(sp, str_sub(species, 1, 2)),
    sp = tolower(sp),
    latin = str_c(genus, species, sep = " "),
    latin = to_any_case(latin)
  ) %>% 
  select(sp = spcode, genus, species, latin, family, trait_group)

@@

```{r, eval = run_scbi}
scbi_2013 <- 
  "https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/raw/master/tree_main_census/data/census-csv-files/scbi.stem2.csv" %>% 
  read_csv() %>% 
  select(
    treeID, stemID, sp, quadrat, gx, gy, dbh, 
    date = ExactDate, codes, status
  )

@@

```{r, eval = run_bw}
census_df1 <- bw_2008
# we need to filter out the resprouts
census_df2 <- bw_2014 %>% 
  filter(!str_detect(codes, 'R'))
id <- "treeID"

@@

```{r, eval = run_scbi}
census_df1 <- scbi_2013
census_df2 <- scbi_2018
id <- "stemID"

@@

```{r, eval = run_scbi & run_bw}
growth_df <- bind_rows(
  bw_growth_df %>% select(growth) %>% mutate(site = "bw"),
  scbi_growth_df %>% select(growth) %>% mutate(site = "scbi")
)
ggplot(growth_df, aes(x = growth, y = ..density.., fill = site)) +
  geom_histogram(alpha = 0.5, position = "identity", binwidth = 0.05) +
  labs(x = "Average annual growth in dbh (cm per yr)") +
  coord_cartesian(xlim = c(-0.5, 1))
```

@@

```{r, eval = run_bw}
# Bigwoods study region boundary polygon
bw_boundary <- bigwoods_study_region %>% 
  sf_polygon()

@@

```{r, eval = run_scbi}
# Bigwoods study region boundary polygon
scbi_boundary <- tibble(x = c(0,400,400,0,0), y = c(0,0,640,640,0)) %>% 
  sf_polygon()

@@

```{r, eval = run_bw}
set.seed(76)
bw_cv_grid <- spatialBlock(
  speciesData = bw_growth_df, theRange = 100, verbose = FALSE,
  # Some guess work in figuring this out:
  k = 28, xOffset = 0.5, yOffset = 0
)

@@

```{r, eval = run_scbi}
scbi_cv_grid <- spatialBlock(
  speciesData = scbi_growth_df, theRange = 100, k = 28, yOffset = 0.9999, verbose = FALSE
)

@@

```{r, eval = run_bw}
bw_specs <- bw_growth_df %>% 
  get_model_specs(model_number = 3, species_notion = 'trait_group')
bw_specs
```

@@

```{r, eval = run_bw}
tic()
focal_vs_comp_bw <- bw_growth_df %>% 
  create_focal_vs_comp(max_dist, model_specs = bw_specs, cv_grid = bw_cv_grid, id = "treeID")
toc()
```

```{r, eval = run_bw}
glimpse(focal_vs_comp_bw)
```

@@

```{r, eval = run_scbi}
scbi_specs <- scbi_growth_df %>% 
  get_model_specs(model_number = 3, species_notion = 'sp')
scbi_specs
```

```{r, eval = run_scbi}
tic()
focal_vs_comp_scbi <- scbi_growth_df %>% 
  create_focal_vs_comp(max_dist, model_specs = scbi_specs, cv_grid = scbi_cv_grid, id = "stemID")
toc()
```

```{r, eval = run_scbi}
glimpse(focal_vs_comp_scbi)
```

@@

```{r, eval = run_bw}
tic()
bw_fit_model <- focal_vs_comp_bw %>% 
  fit_bayesian_model(model_specs = bw_specs)
toc()
```

@@

```{r, eval = run_bw}
bw_growth_df <- focal_vs_comp_bw %>% 
  predict_bayesian_model(model_specs = bw_specs, posterior_param = bw_fit_model) %>% 
  right_join(bw_growth_df, by = c("focal_ID" = "treeID"))

@@

```{r, eval = run_scbi}
tic()
scbi_fit_model <- focal_vs_comp_scbi %>% 
  fit_bayesian_model(model_specs = scbi_specs)
toc()
```

```{r, eval = run_scbi}
scbi_growth_df <- focal_vs_comp_scbi %>% 
  predict_bayesian_model(model_specs = scbi_specs, posterior_param = scbi_fit_model) %>% 
  right_join(scbi_growth_df, by = c("focal_ID" = "stemID"))

@@

```{r, eval = run_scbi}
tic()
scbi_cv_predict <- focal_vs_comp_scbi %>%
  run_cv(model_specs = scbi_specs, max_dist = max_dist, cv_grid = scbi_cv_grid, all_folds = FALSE)
toc()
```

@@

```{r, eval = run_scbi}
scbi_cv_predict %>%
  inner_join(scbi_growth_df, by = 'focal_ID', suffix = c('_cv','')) %>%
  summarise(
    rmse_cv = sqrt(mean((growth - growth_hat_cv)^2)),
    rmse = sqrt(mean((growth - growth_hat)^2)), 
    n = n() 
  )
```

@@

```{r, eval = run_bw}
tic()
bw_cv_predict <- focal_vs_comp_bw %>%
  run_cv(model_specs = bw_specs, max_dist = max_dist, cv_grid = bw_cv_grid)
toc()
```

@@

```{r, eval = run_bw}
bw_growth_df %>% 
  left_join(bw_cv_predict, by = 'focal_ID', suffix = c('', '_cv')) %>% 
  mutate(has_cv = !is.na(growth_hat_cv)) %>% 
  st_as_sf() %>% 
  ggplot() +
  geom_sf(size = 0.1, aes(col = has_cv))
```

>>>>

```{r, eval = run_bw}
# Big Woods
# Read in census data from 2008 & 2014
bw_2008 <- 
  "https://deepblue.lib.umich.edu/data/downloads/z603qx485" %>% 
  read_delim(delim = "\t") %>% 
  mutate(spcode = to_any_case(spcode)) %>%
  select(
    treeID = treeid, stemID = stemtag, sp = spcode, quadrat, gx, gy, dbh, 
    date, codes
  )
bw_2014 <- 
  "https://deepblue.lib.umich.edu/data/downloads/1831ck00f" %>% 
  read_delim(delim = "\t") %>% 
  mutate(spcode = to_any_case(spcode)) %>%
  select(
    treeID = treeid, stemID = stemtag, sp = spcode, quadrat, gx, gy, dbh, 
    date, codes
  )

# Read in grouping classification data
bw_species <- 
  "https://deepblue.lib.umich.edu/data/downloads/000000086" %>% 
  read_delim(delim = "\t") %>% 
  # convert all to snake case:
  mutate_at(c("species", "genus", "family", "idlevel", "spcode"), to_any_case) %>% 
  # join trait group
  left_join(families, by = c("spcode", "family")) %>% 
  mutate(
    sp = str_sub(genus, 1, 2), 
    sp = str_c(sp, str_sub(species, 1, 2)),
    sp = tolower(sp),
    latin = str_c(genus, species, sep = " "),
    latin = to_any_case(latin)
  ) %>% 
  select(sp = spcode, genus, species, latin, family, trait_group)

@@

```{r, eval = run_scbi}
# SCBI
scbi_2013 <- 
  "https://github.com/SCBI-ForestGEO/SCBI-ForestGEO-Data/raw/master/tree_main_census/data/census-csv-files/scbi.stem2.csv" %>% 
  read_csv() %>% 
  select(
    treeID, stemID, sp, quadrat, gx, gy, dbh, 
    date = ExactDate, codes, status
  )

@@

```{r, eval = run_bw}
# Big Woods
census_df1 <- bw_2008
# we need to filter out the resprouts
census_df2 <- bw_2014 %>% 
  filter(!str_detect(codes, 'R'))
id <- "treeID"

@@

```{r, eval = run_scbi}
# SCBI
census_df1 <- scbi_2013
census_df2 <- scbi_2018
id <- "stemID"

@@

```{r, eval = run_scbi & run_bw}
# Both Big Woods & SCBI
growth_df <- bind_rows(
  bw_growth_df %>% select(growth) %>% mutate(site = "bw"),
  scbi_growth_df %>% select(growth) %>% mutate(site = "scbi")
)
ggplot(growth_df, aes(x = growth, y = ..density.., fill = site)) +
  geom_histogram(alpha = 0.5, position = "identity", binwidth = 0.05) +
  labs(x = "Average annual growth in dbh (cm per yr)") +
  coord_cartesian(xlim = c(-0.5, 1))
```

@@

```{r, eval = run_bw}
# Bigwoods
# Study region boundary polygon
bw_boundary <- bigwoods_study_region %>% 
  sf_polygon()

@@

```{r, eval = run_scbi}
# SCBI
# Study region boundary polygon
scbi_boundary <- tibble(x = c(0,400,400,0,0), y = c(0,0,640,640,0)) %>% 
  sf_polygon()

@@

```{r, eval = run_bw}
# Big Woods
set.seed(76)
bw_cv_grid <- spatialBlock(
  speciesData = bw_growth_df, theRange = 100, verbose = FALSE,
  # Some guess work in figuring this out:
  k = 28, xOffset = 0.5, yOffset = 0
)

@@

```{r, eval = run_scbi}
# SCBI
scbi_cv_grid <- spatialBlock(
  speciesData = scbi_growth_df, theRange = 100, k = 28, yOffset = 0.9999, verbose = FALSE
)

@@

```{r, eval = run_bw}
# Big Woods
bw_specs <- bw_growth_df %>% 
  get_model_specs(model_number = 3, species_notion = 'trait_group')
bw_specs
```

@@

```{r, eval = run_bw}
# Big Woods
if (!file.exists("focal_vs_comp_bw.Rdata")) {
  tic()
  focal_vs_comp_bw <- bw_growth_df %>% 
    create_focal_vs_comp(max_dist, model_specs = bw_specs, cv_grid = bw_cv_grid, id = "treeID")
  toc()
  save(focal_vs_comp_bw, file = "focal_vs_comp_bw.Rdata")
} else {
  load("focal_vs_comp_bw.Rdata")
}
```

```{r, eval = run_bw}
# Big Woods
glimpse(focal_vs_comp_bw)
```

@@

```{r, eval = run_scbi}
# SCBI
scbi_specs <- scbi_growth_df %>% 
  get_model_specs(model_number = 3, species_notion = 'sp')
scbi_specs
```

```{r, eval = run_scbi}
# SCBI
tic()
focal_vs_comp_scbi <- scbi_growth_df %>% 
  create_focal_vs_comp(max_dist, model_specs = scbi_specs, cv_grid = scbi_cv_grid, id = "stemID")
toc()
```

```{r, eval = run_scbi}
# SCBI
glimpse(focal_vs_comp_scbi)
```

@@

```{r, eval = run_bw}
# Big Woods
tic()
bw_fit_model <- focal_vs_comp_bw %>% 
  fit_bayesian_model(model_specs = bw_specs)
toc()
```

@@

```{r, eval = run_bw}
# Big Woods
bw_growth_df <- focal_vs_comp_bw %>% 
  predict_bayesian_model(model_specs = bw_specs, posterior_param = bw_fit_model) %>% 
  right_join(bw_growth_df, by = c("focal_ID" = "treeID"))

@@

```{r, eval = run_scbi}
# SCBI
tic()
scbi_fit_model <- focal_vs_comp_scbi %>% 
  fit_bayesian_model(model_specs = scbi_specs)
toc()
```

```{r, eval = run_scbi}
# SCBI
scbi_growth_df <- focal_vs_comp_scbi %>% 
  predict_bayesian_model(model_specs = scbi_specs, posterior_param = scbi_fit_model) %>% 
  right_join(scbi_growth_df, by = c("focal_ID" = "stemID"))

@@

```{r, eval = run_scbi}
# SCBI
tic()
scbi_cv_predict <- focal_vs_comp_scbi %>%
  run_cv(model_specs = scbi_specs, max_dist = max_dist, cv_grid = scbi_cv_grid, all_folds = FALSE)
toc()
```

@@

```{r, eval = run_scbi}
# SCBI
scbi_cv_predict %>%
  inner_join(scbi_growth_df, by = 'focal_ID', suffix = c('_cv','')) %>%
  summarise(
    rmse_cv = sqrt(mean((growth - growth_hat_cv)^2)),
    rmse = sqrt(mean((growth - growth_hat)^2)), 
    n = n() 
  )
```

@@

```{r, eval = run_bw}
# Big Woods
tic()
bw_cv_predict <- focal_vs_comp_bw %>%
  run_cv(model_specs = bw_specs, max_dist = max_dist, cv_grid = bw_cv_grid)
toc()
```

@@

```{r, eval = run_bw}
# Big Woods
bw_growth_df %>% 
  left_join(bw_cv_predict, by = 'focal_ID', suffix = c('', '_cv')) %>% 
  mutate(has_cv = !is.na(growth_hat_cv)) %>% 
  st_as_sf() %>% 
  ggplot() +
  geom_sf(size = 0.1, aes(col = has_cv))
```
